\section{Discussion}
This experiment aimed and suceeded to create and analyze entangled photon pairs using the four Bell states (\( |\phi^+\rangle \), \( |\phi^-\rangle \), \( |\psi^+\rangle \), and \( |\psi^-\rangle \)). Additionally, the violation of Bell's inequality through the CHSH parameter demonstrates the presence of non-classical correlations.

For \( |\phi^+\rangle \), we achieved the highest fidelity 0.8817 and a purity of 0.7973, which suggests that this state was prepared with great accuracy.
\( |\psi^+\rangle \) also showed high fidelity 0.8710 and a similarly strong purity 0.7837. The partial transpose of the density matrix of all the states had negative eigenvalues which show that they are entangled states.
On the other hand, \( |\phi^-\rangle \) and \( |\psi^-\rangle \) had slightly lower fidelities (0.8085 and 0.8039, respectively) and purities (0.7080 and 0.7838). This is likely due to small imperfections in the alignments using waveplates or the crystals. 

The CHSH parameter we measured in the experiment is \( S = 2.385 \pm 0.015 \). This exceeds the classical limit of \( S = 2 \),which violates Bell’s inequality. This validates the presence of entanglement and non-classical correlations in the photon pairs.

Some of the key factors that could explain the lower fidelities and purities observed include:

\begin{itemize}
    \item Photon loss in the optical components and during coupling to the fibers.
    \item Limited quantum efficiency of the Spontaneous Parametric Down Conversion process.
    \item Detector inefficiencies, such as dead time or recovery time or shot noise, reducing accurate photon detection.
    \item Misalignments in the waveplates or crystals, introducing phase and polarization errors.
\end{itemize}

High-quality Bell states  are essential for a range of quantum technologies, including quantum communication protocols like quantum key distribution and teleportation. Violating Bell's inequality is one of the foundational tests of quantum mechanics. Entangled photon pairs have correlations which rule out local realistic theories such as hidden variable theories that allow for a local and realistic description of the universe.
